---
title: "Global Terrorist Activities Time Series Analysis and Iraq Spatial Case Study"
author: "Yuan Gao, Qi Wang, Yizheng Wang, Darien Zhang"
date: "4/21/2017"
output: pdf_document
---

## Introduction

This project aims to explore and analyze the dataset from Global Terrorism Database that contains details on every terrorist activity since 1970. There are two parts of the project. The first part is creating ARIMA models for aggregated count data for terrorist activities in each region of the world and forecasting for the future years. The second part is 

## Data

The Global Terrorism Database (GTD) is maintained by the National Consortium for the Study of Terrorism and Responses to Terrorism (START). Variables include incident date, incident information, incident location, attack information, weapon information, target information, and perpetrator information. There are 156,772 incidents ranging from 1970 to 2015.



## Part I: Regional Time Series Analysis

The first part of the project is to build ARIMA models for all twelve regions on aggregated count of the incidents and perform forecasts for the next several years. We performed the analysis on both yearly and monthly aggregated count data.

There is obvious drift in the original time series. For fulfilling stationarity, we have visualize the difference of the original data. Corresponding ACF and PACF are also plotted for modeling.

A periodogram can be used to identify the dominant periods (or frequencies) of a time series.  It is a helpful tool in spectral analysis to detect cyclical behavior. Typically the time series are observed in time domain, but it can also be transformed to the frequency domain by Fourier Transforamtion. If there is significant peak at certain frequency, there may be significant periodicity in the series.

Based on observing the ACF, PACF and periodogram, by grid searching on a small range, here we have presented the best models. It turns out to have relatively low AIC values and the residuals are not autocorrelated. Thus, the models are good enough for us to use for forecast.

Similarly, we have also constructed models on the monthly data.


\newpage
## Appendix

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# Load packages
library(dplyr)
library(forecast)
library(ggplot2)
library(modelr)
library(tidyr)
```

```{r,echo = F,fig.height=8,fig.width=7}
#Load data
data = read.csv('globalterrorismdb_0616dist.csv', header = T)
regions = data[,c('region','region_txt')]
reg_uniq = unique(regions) 
reg_uniq = reg_uniq[order(reg_uniq$region),]
n = nrow(reg_uniq)

# Create yearly and monthly data
for (i in 1:n){
  var = paste0("data",i,"_count")
  var1 = paste0("data",i,"_month")
  assign(var, data %>% filter(region == i) %>% group_by(iyear) %>% summarise(count = n()))
  assign(var1, data %>% filter(region == i) %>% group_by(iyear, imonth) %>% summarise(count = n()))
}

for (i in 1:n){
  use = get(paste0("data",i,"_month"))
  use = use[which(use$imonth != 0),]
  assign(paste0("data",i,"_month"),use)
}

for(i in 1:n){
  d = get(paste0("data",i,"_month"))
  assign(paste0("ts",i),ts(d$count, frequency = 12, start = c(1970,1)))
}

```

```{r,echo = F,fig.height=6.5,fig.width=6, fig.align = 'center'}
par(mfrow = c(4,3),mar=c(2,2,2,1))

for(i in 1:n){
  plot(get(paste0("data",i,"_count")), type = 'l', main = reg_uniq[i,'region_txt'])
}


```

\begin{center}
Yearly Count
\end{center}



```{r,eval = F, fig.height=8,fig.width=7, echo = F}
par(mfrow = c(3,4))
for (i in 1:12){
  var = paste0("fit",i)
  dat = get(paste0("data",i,"_count"))
  da = ts(dat$count, start = 1970)
  assign(var, auto.arima(da, stepwise = FALSE, seasonal = TRUE))
  this = get(paste0("fit",i))
  plot(forecast(this,h=5), main = reg_uniq[i,'region_txt'])
  var1 = paste0("residual",i)
  assign(var1, this$residuals)
}

par(mfrow=c(1,3), mar=c(2,4,6,2))
for (i in 1:n){
  res = get(paste0("residual",i))
  plot(res,main=reg_uniq[i,'region_txt'])
  acf(res)
  pacf(res)
}


par(mfrow = c(3,4))
for (i in 1:12){
  var = paste0("fit_month",i)
  dat = get(paste0("ts",i))
  assign(var, auto.arima(dat, stepwise = FALSE, seasonal = TRUE))
  this = get(paste0("fit_month",i))
  plot(forecast(this,h=50), main = reg_uniq[i,'region_txt'])
  var1 = paste0("residual_month",i)
  assign(var1, this$residuals)
}

par(mfrow=c(1,3), mar=c(2,4,6,2))
for (i in 1:n){
  res = get(paste0("residual_month",i))
  plot(res,main=reg_uniq[i,'region_txt'])
  acf(res)
  pacf(res)
}
```

```{r,echo = F}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

```{r, eval = F, echo = F}

for(i in 1:12){
  data_in_use = get(paste0("data",i,"_count"))
  l = lm(count~iyear, data = data_in_use)
  l2 = lm(count ~ iyear + I(iyear^2), data=data_in_use)
  
  d = data_in_use %>%
    add_predictions(l, var="l") %>%
    add_predictions(l2, var="l2")
  
  var = paste0("p",i)
  assign(var,ggplot(data_in_use, aes(x=iyear, y=count)) + 
           geom_line() + 
           geom_point() +
           geom_line(data=d, aes(x=iyear, y=l), color="red", size=1.2) +
           geom_line(data=d, aes(x=iyear, y=l2), color="blue", size=1.2)+
           ggtitle(reg_uniq[i,'region_txt']))
  
  d = data_in_use %>%
    add_residuals(l, var="resid_l") %>%
    add_residuals(l2, var="resid_q")
  
  var = paste0("r",i)
  
  assign(var,ggplot(gather(d, type, residual, -(iyear:count)), aes(x=iyear, y=residual, color=type)) + 
           geom_point() +
           geom_line() +
           facet_wrap(~type, nrow=2)+
           ggtitle(reg_uniq[i,'region_txt']))
}
multiplot(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,p12, cols = 4)
multiplot(r1,r2,r3,r4,r5,r6,r7,r8,r9,r10,r11,r12, cols = 4)


```







```{r, echo = F, fig.height=6.5,fig.width=6.5}
# ACF and PACF for yearly data

par(mfrow = c(4,3),mar=c(1,2,3,1))

for(i in 1:n){
  var = paste0("diff",i)
  data_c = get(paste0("data",i,"_count"))
  assign(var, diff(data_c$count))
  data_diff = get(paste0("diff",i))
  plot(data_diff, type = 'l', main = reg_uniq[i,'region_txt'])

}
```

\begin{center}
Differenced Data
\end{center}

```{r, echo = F, fig.height=6.5,fig.width=6.5}
par(mfrow = c(4,3),mar=c(1,2,3,1))

for(i in 1:n){
  data_diff = get(paste0("diff",i))
  acf(data_diff,main = reg_uniq[i,'region_txt'])
}

```

\begin{center}
acf
\end{center}

```{r, echo = F, fig.height=6.5,fig.width=6.5}
par(mfrow = c(4,3),mar=c(1,2,3,1))

for(i in 1:n){
  data_diff = get(paste0("diff",i))
  pacf(data_diff,main = reg_uniq[i,'region_txt'])
}
```

\begin{center}
pacf
\end{center}

```{r,fig.height=6.5,fig.width=6, echo = F, eval = F}
# Periodogram for yearly data
par(mfrow = c(4,3),mar=c(4,2,1.5,1))
for(i in 1:n){
  data_diff = get(paste0("diff",i))
  var = paste0("perio",i)
  assign(var,spec.pgram(data_diff,taper=0,log="no",main=reg_uniq[i,'region_txt']))
}

spec.pgram(diff11,taper=0,log="no",main=reg_uniq[11,'region_txt'])
```

\begin{center}
Periodogram
\end{center}


```{r, fig.height=6.5,fig.width=6, echo = F}
# (S)ARIMA modeling for 12 regions

par(mfrow = c(4,3),mar=c(4,2,1.5,1))

# North America
testarma=Arima(data1_count$count,order=c(1,1,1),seasonal=list(order=c(0,0,1),period=6))
testarma_res1 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=5)
plot.forecast(pmdarmaforecast,main=reg_uniq[1,'region_txt'])


testarma=Arima(data2_count$count,order=c(0,1,1),seasonal=list(order=c(1,0,1),period=2))
testarma_res2 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=5)
plot.forecast(pmdarmaforecast,main=reg_uniq[2,'region_txt'])

testarma=Arima(data3_count$count,order=c(5,1,1),seasonal=list(order=c(2,0,1),period=4))
testarma_res3 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=5)
plot.forecast(pmdarmaforecast,main=reg_uniq[3,'region_txt'])

testarma=Arima(data4_count$count,order=c(3,1,3))
testarma_res4 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=5)
plot.forecast(pmdarmaforecast,main=reg_uniq[4,'region_txt'])

testarma=Arima(data5_count$count,order=c(1,1,1),seasonal=list(order=c(2,0,1),period=4))
testarma_res5 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=5)
plot.forecast(pmdarmaforecast,main=reg_uniq[5,'region_txt'])

testarma=Arima(data6_count$count,order=c(0,1,2),seasonal=list(order=c(2,0,1),period=6))
testarma_res6 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=5)
plot.forecast(pmdarmaforecast,main=reg_uniq[6,'region_txt'])

testarma=Arima(data7_count$count,order=c(1,1,1),seasonal=list(order=c(2,0,1),period=2))
testarma_res7 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=5)
plot.forecast(pmdarmaforecast,main=reg_uniq[7,'region_txt'])

testarma=Arima(data8_count$count,order=c(5,2,3))
testarma_res8 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=5)
plot.forecast(pmdarmaforecast,main=reg_uniq[8,'region_txt'])

testarma=Arima(data9_count$count,order=c(3,1,2),seasonal=list(order=c(0,0,1),period=2))
testarma_res9 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=5)
plot.forecast(pmdarmaforecast,main=reg_uniq[9,'region_txt'])

testarma=Arima(data10_count$count,order=c(5,1,5),seasonal=list(order=c(3,0,1),period=4))
testarma_res10 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=5)
plot.forecast(pmdarmaforecast,main=reg_uniq[10,'region_txt'])

testarma=Arima(data11_count$count,order=c(5,1,3),seasonal=list(order=c(2,0,1),period=2))
testarma_res11 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=5)
plot.forecast(pmdarmaforecast,main=reg_uniq[11,'region_txt'])

testarma=Arima(data12_count$count,order=c(2,1,3))
testarma_res12 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=5)
plot.forecast(pmdarmaforecast,main=reg_uniq[12,'region_txt'])


```

\begin{center}
Yearly Forecast
\end{center}

```{r,fig.height=9,fig.width=7,echo=FALSE}
# Checking modeling residuals
par(mfrow=c(6,3), mar=c(2,4,6,2))
for (i in 1:n){
  testarma_res = get(paste0("testarma_res",i))
  plot(testarma_res,main=reg_uniq[i,'region_txt'])
  acf(testarma_res)
  pacf(testarma_res)
}
```

\begin{center}
Residuals
\end{center}






```{r,echo = F,fig.height=6.5,fig.width=6, fig.align = 'center'}
par(mfrow = c(4,3),mar=c(4,2,1.5,1))
for(i in 1:n){
  plot(get(paste0("ts",i)),  main = reg_uniq[i,'region_txt'], ylab = "count")
}



```

\begin{center}
Monthly Original Data
\end{center}

```{r, fig.height=8,fig.width=7, echo=FALSE}
# ACF and PACF for monthly data

par(mfrow = c(4,3),mar=c(1,2,3,1))

for(i in 1:n){
  var = paste0("diff",i)
  data_c = get(paste0("data",i,"_month"))
  assign(var, diff(data_c$count))
  data_diff = get(paste0("diff",i))
  plot(data_diff, type = 'l', main = reg_uniq[i,'region_txt'])

}
```
\begin{center}
Differenced Data
\end{center}

```{r, fig.height=8,fig.width=7, echo=F}
par(mfrow = c(4,3),mar=c(1,2,3,1))
for(i in 1:n){
  data_diff = get(paste0("diff",i))
  acf(data_diff,main = reg_uniq[i,'region_txt'])
}
```

\begin{center}
acf
\end{center}

```{r, fig.height=8,fig.width=7, echo=F}
par(mfrow = c(4,3),mar=c(1,2,3,1))
for(i in 1:n){
  data_diff = get(paste0("diff",i))
  pacf(data_diff,main = reg_uniq[i,'region_txt'])
}
```

\begin{center}
pacf
\end{center}

```{r, fig.height=8,fig.width=7, echo=F}
# Periodogram for monthly data
par(mfrow = c(4,3))
for(i in 1:n){
  data_diff = get(paste0("diff",i))
  var = paste0("perio",i)
  assign(var,spec.pgram(data_diff,taper=0,log="no",main=reg_uniq[i,'region_txt']))
}

```

\begin{center}
Periodogram
\end{center}

```{r, fig.height=8,fig.width=7, echo = F}
# (S)ARIMA modeling for 12 regions

par(mfrow = c(4,3))

# North America
testarma=Arima(ts1,order=c(5,1,1))
testarma_res1 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=20)
plot.forecast(pmdarmaforecast,main=reg_uniq[1,'region_txt'])


testarma=Arima(ts2,order=c(5,1,5),seasonal=list(order=c(1,0,1),period=2))
testarma_res2 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=20)
plot.forecast(pmdarmaforecast,main=reg_uniq[2,'region_txt'])


testarma=Arima(ts3,order=c(7,1,5),seasonal=list(order=c(2,0,1),period=4))
testarma_res3 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=20)
plot.forecast(pmdarmaforecast,main=reg_uniq[3,'region_txt'])

testarma=Arima(ts4,order=c(5,1,2),seasonal=list(order=c(1,0,1),period=3))
testarma_res4 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=20)
plot.forecast(pmdarmaforecast,main=reg_uniq[4,'region_txt'])

testarma=Arima(ts5,order=c(6,1,1))
testarma_res5 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=20)
plot.forecast(pmdarmaforecast,main=reg_uniq[5,'region_txt'])

testarma=Arima(ts6,order=c(5,1,5),seasonal=list(order=c(2,0,1),period=2))
testarma_res6 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=20)
plot.forecast(pmdarmaforecast,main=reg_uniq[6,'region_txt'])

testarma=Arima(ts7,order=c(5,1,5),seasonal=list(order=c(1,0,1),period=3))
testarma_res7 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=20)
plot.forecast(pmdarmaforecast,main=reg_uniq[7,'region_txt'])

testarma=Arima(ts8,order=c(3,1,5),seasonal=list(order=c(1,0,1),period=3))
testarma_res8 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=20)
plot.forecast(pmdarmaforecast,main=reg_uniq[8,'region_txt'])

testarma=Arima(ts9,order=c(2,1,5),seasonal=list(order=c(0,0,1),period=3))
testarma_res9 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=20)
plot.forecast(pmdarmaforecast,main=reg_uniq[9,'region_txt'])

testarma=Arima(ts10,order=c(5,1,5),seasonal=list(order=c(1,0,1),period=2))
testarma_res10 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=20)
plot.forecast(pmdarmaforecast,main=reg_uniq[10,'region_txt'])

testarma=Arima(ts11,order=c(3,1,3),seasonal=list(order=c(1,0,1),period=3))
testarma_res11 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=20)
plot.forecast(pmdarmaforecast,main=reg_uniq[11,'region_txt'])

testarma=Arima(ts12,order=c(3,1,2),seasonal=list(order=c(0,0,1),period=2))
testarma_res12 = testarma$residuals
pmdarmaforecast=forecast.Arima(testarma,h=20)
plot.forecast(pmdarmaforecast,main=reg_uniq[12,'region_txt'])
```

\begin{center}
Forecast - Monthly
\end{center}

```{r, fig.height=9,fig.width=7, echo = F}
# Checking modeling residuals
par(mfrow=c(6,3), mar=c(2,4,6,2))
for (i in 1:n){
  testarma_res = get(paste0("testarma_res",i))
  plot(testarma_res,main=reg_uniq[i,'region_txt'])
  acf(testarma_res)
  pacf(testarma_res)
}
```

\begin{center}
Residuals
\end{center}